### 2018-04-26 doc/tlla2018/main.tex
minor changes

\documentclass[draft,submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{Linearity and TLLA 2018}
\usepackage[british]{babel}
\usepackage[all]{foreign}
\usepackage[fleqn]{amsmath}
\usepackage{amsthm}
\allowdisplaybreaks
\usepackage[inline]{enumitem}
\usepackage{breakurl} % Not needed if you use pdflatex only.
\usepackage{underscore} % Only needed if you use pdflatex.
\usepackage{textcomp,textgreek,upgreek}
\usepackage{nameref,hyperref}
\usepackage[capitalise,nameinlink,noabbrev]{cleveref}
\hypersetup{hidelinks,final}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}
\input{preamble-typing}
%\usepackage[parfill]{parskip}
\usepackage{more-space}
\usepackage[sectionbib,square,numbers,sort&compress]{natbib}

\title{Taking Linear Logic Apart}
\author{%
  Wen Kokke
  \institute{University of Edinburgh\\ Edinburgh, Scotland}
  \email{wen.kokke@ed.ac.uk}
  \and
  Fabrizio Montesi
  \institute{University of Southern Denmark\\ Odense, Denmark}
  \email{fmontesi@imada.sdu.dk}
  \and
  Marco Peressotti
  \institute{University of Southern Denmark\\ Odense, Denmark}
  \email{peressotti@imada.sdu.dk}}
\def\titlerunning{Taking Linear Logic Apart}
\def\authorrunning{W.\ Kokke, F.\ Montesi, and M.\ Peressotti}

\begin{document}
\maketitle

\begin{abstract}
  Process calculi based in logic, such as \piDILL and \cp, provide a foundation
  for deadlock-free concurrent programming.
  However, there is a mismatch between the structures of operators used as proof
  terms in previous work, and the term constructs of the standard \textpi-calculus.
  We introduce the Hypersequent Classical Processes (\hcp), which addresses this
  mismatch.
  The key insight is to register parallelism in the typing judgements using
  hypersequents, a technique from logic which generalises judgements from one
  sequent to many.
  This allows us to take apart the term constructs used in Classical Processes
  (\cp) to more closely match those of the standard \textpi-calculus.
  We prove that \hcp enjoys subject reduction and progress, and prove several
  properties relating it back to \cp.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Classical Processes (\cp) \cite{wadler2012} is a process calculus inspired by the
correspondence between the session-typed \textpi-calculus and linear logic
\cite{caires2010}, where processes correspond to proofs, session types
(communication protocols) to propositions, and communication to cut elimination.
Bridging process languages to linear logic paves the way for applying methods
developed in one field to the other.
For instance, the proof theory of linear logic can be used to guarantee progress
for processes~\cite{caires2010,wadler2012}.

The main attraction of \cp is that its semantics are \emph{prescribed} by the
cut elimination procedure of Classical Linear Logic (CLL).
This permits us to reuse the metatheory of linear logic ``as is'' to reason about
the behaviour of processes.
However, there is a mismatch between the structure of the proof terms of CLL and
the term constructs of the standard \textpi-calculus~\cite{milner1992a,milner1992b}.
For instance, the term for output of a linear name is
$\tm{\cpSend{x}{y}{P}{Q}}$, which is read ``send $y$ over $x$ and proceed as $P$
in parallel to $Q$''.
Note that this is a single term constructor, which takes all four arguments at
the same time.
This is caused by directly adopting the $(\tens)$-rule from CLL as the process
calculus construct for sending: the $(\tens)$-rule has two premises (corresponding to $\tm P$ and $\tm Q$ in the output term), and checks
that they share no resources (in the output term, $\tm x$ can be used only by $\tm P$, and $\tm y$ can be used only by $\tm Q$).

There is no independent parallel term $\tm{(\piPar{P}{Q})}$ in the grammar of
\cp terms.
Instead, the parallel composition shows up in any term which corresponds to a
binary rule.
Even if we were to add an independent parallel composition via the
\textsc{Mix}-rule, as suggested in the original presentation of
\cp~\cite{wadler2012}, there would be no way to allow the composed process
$\tm{P}$ and $\tm{Q}$ to communicate as in the standard \textpi-calculus, as
there is no independent name restriction either!
Instead, synchronisation is governed by the ``cut'' operator
$\tm{\piNew{x}{(\piPar{P}{Q})}}$, which composes $\tm{P}$ and $\tm{Q}$, enabling
them to communicate along $\tm{x}$.
Worse, if we naively added an independent parallel composition as well as a
name restriction, using the rules shown below, we would lose cut elimination,
and therefore deadlock-freedom!
\begin{center}
  \begin{prooftree*}
    \AXC{$\seq[ P ]{ \Gamma }$}
    \AXC{$\seq[ Q ]{ \Delta }$}
    \NOM{Mix}
    \BIC{$\seq[ \piPar{P}{Q} ]{ \Gamma , \Delta }$}
  \end{prooftree*}
  \begin{prooftree*}
    \AXC{$\seq[ P ]{ \Gamma, \tmty{x}{A}, \tmty{y}{A^\bot} }$}
    \NOM{``Cut''}
    \UIC{$\seq[ \piNew{xy}{P} ]{ \Gamma }$}
  \end{prooftree*}
\end{center}

This syntactic mismatch has an effect on the semantics as well.
For instance, the $\beta$-reduction for output and input in \cp is
\(
\reducesto
{\cpCut{x}{\cpSend{x}{y}{P}{Q}}{\cpRecv{x}{y}{R}}}
{\cpCut{y}{P}{\cpCut{x}{Q}{R}}}
\).
Here, the parallel composition $\tm{(\piPar{P}{Q})}$ is of no relevance to this
communication, yet the rule needs to inspect it to be able to nest the name
restrictions appropriately in the resulting term.

In this paper, we introduce the Hypersequent Classical Processes (\hcp), which
addresses this mismatch.
The key insight is to register parallelism in the typing judgements using
hypersequents, a technique from logic which generalises judgements from one
sequent to many.
This allows us to take apart the term constructs used in Classical Processes
(\cp) to more closely match those of the standard \textpi-calculus.
We proceed as follows.
We start by introducing \cp (\cref{sec:cp}).
Then, we introduce \hcp and prove it enjoys subject reduction and progress
(\cref{sec:hcp}).
We prove several properties relating the two.
Specifically, we prove that every \cp process is an \hcp process, that every
\hcp process corresponds to several entangled \cp processes, and that \hcp
supports the same communication protocols as \cp, and no more
(\cref{sec:cp2hcp}).
Finally, we discuss related work and conclude
(\cref{sec:related-work,sec:conclusions}).

\section{Classical Processes}
\label{sec:cp}

In order to keep the discussion of \hcp in \cref{sec:hcp} simple, we restrict
ourselves to the multiplicative-applicative subset of \cp.
We foresee no problems in extending the proofs in \cref{sec:hcp} to cover the
remaining features of \cp (polymophism and the exponentials).

The term language of \cp is a variant of the \textpi-calculus.
The variables $\tm{x}$, $\tm{y}$, and $\tm{z}$ range over channel names.
The construct $\tm{\cpLink{x}{y}}$ links two
channels~\cite{sangiorgi1996,boreale1998}, forwarding messages received on
$\tm{x}$ to $\tm{y}$ and vice versa.
The construct $\tm{\cpCut{x}{P}{Q}}$ creates a new channel $\tm{x}$, and composes
two processes, $\tm{P}$ and $\tm{Q}$, which communicate on $\tm{x}$, in parallel.
Therefore, in $\tm{\cpCut{x}{P}{Q}}$ the name $\tm{x}$ is bound in both $\tm{P}$
and $\tm{Q}$.
In $\tm{\cpRecv{x}{y}{P}}$ and $\tm{\cpSend{x}{y}{P}{Q}}$, round brackets denote
input, square brackets denote output.
We use bound output~\cite{sangiorgi1996}, meaning that both input and output
bind a new name.
In $\tm{\cpRecv{x}{y}{P}}$ the new name $\tm{y}$ is bound in $\tm{P}$.
In $\tm{\cpSend{x}{y}{P}{Q}}$, the new name $\tm{y}$ is only bound in $\tm{P}$,
while $\tm{x}$ is only bound in $\tm{Q}$.
\begin{definition}[Terms]\label{def:cp-terms}  Process terms are given by the following grammar:
  \[
    \begin{array}{rllrll}
      \tm{P}, \tm{Q}, \tm{R}
           :=& \tm{\cpLink{x}{y}}       &\text{link}
      &  \mid& \tm{\cpCut{x}{P}{Q}}     &\text{parallel composition, ``cut''}
      \\ \mid& \tm{\cpSend{x}{y}{P}{Q}} &\text{output}
      &  \mid& \tm{\cpRecv{x}{y}{P}}    &\text{input}
      \\ \mid& \tm{\cpHalt{x}}          &\text{halt}
      &  \mid& \tm{\cpWait{x}{P}}       &\text{wait}
      \\ \mid& \tm{\cpInl{x}{P}}        &\text{select left choice}
      &  \mid& \tm{\cpInr{x}{P}}        &\text{select right choice}
      \\ \mid& \tm{\cpCase{x}{P}{Q}}    &\text{offer binary choice}
      &  \mid& \tm{\cpAbsurd{x}}        &\text{offer nullary choice}
    \end{array}
  \]
\end{definition}

Terms in \cp are identified up to structural congruence, which states that
links are symmetric, and parallel compositions $\tm{\cpCut{x}{P}{Q}}$ are
associative and commutative.
\begin{definition}[Structural congruence]\label{def:cp-equiv}
  The structural congruence $\equiv$ is the congruence closure over terms which
  satisfies the following additional axioms:
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{llcll}
      \cpEquivLinkComm
      & \tm{\cpLink{x}{y}}
      & \equiv
      & \tm{\cpLink{y}{x}}
      \\
      \cpEquivCutComm
      & \tm{\cpCut{x}{P}{Q}}
      & \equiv
      & \tm{\cpCut{x}{Q}{P}}
      \\
      \cpEquivCutAss1
      & \tm{\cpCut{x}{P}{\cpCut{y}{Q}{R}}}
      & \equiv
      & \tm{\cpCut{y}{\cpCut{x}{P}{Q}}{R}}
      & \text{if }\notFreeIn{x}{R}\text{ and }\notFreeIn{y}{P}
    \end{array}
  \]
\end{definition}

\begin{figure}
 \[
    \begin{array}{llcll}
      \cpRedAxCut1
      & \tm{\cpCut{x}{\cpLink{w}{x}}{P}}
      & \Longrightarrow
      & \tm{\cpSub{w}{x}{P}} 
      \\
      \cpRedBetaTensParr
      & \tm{\cpCut{x}{\cpSend{x}{y}{P}{Q}}{\cpRecv{x}{y}{R}}}
      & \Longrightarrow
      & \tm{\cpCut{y}{P}{\cpCut{x}{Q}{R}}}
      \\
      \cpRedBetaOneBot
      & \tm{\cpCut{x}{\cpHalt{x}}{\cpWait{x}{P}}}
      & \Longrightarrow
      & \tm{P}
      \\
      \cpRedBetaPlusWith1
      & \tm{\cpCut{x}{\cpInl{x}{P}}{\cpCase{x}{Q}{R}}}
      & \Longrightarrow
      & \tm{\cpCut{x}{P}{Q}}
      \\
      \cpRedBetaPlusWith2
      & \tm{\cpCut{x}{\cpInr{x}{P}}{\cpCase{x}{Q}{R}}}
      & \Longrightarrow
      & \tm{\cpCut{x}{P}{R}}
      \\
      \cpRedKappaTens1
      & \tm{\cpCut{x}{\cpSend{y}{z}{P}{Q}}{R}}
      & \Longrightarrow \;
      & \tm{\cpSend{y}{z}{\cpCut{x}{P}{R}}{Q}}
        \quad \text{if} \; \notFreeIn{x}{Q}
      \\
      \cpRedKappaTens2
      & \tm{\cpCut{x}{\cpSend{y}{z}{P}{Q}}{R}}
      & \Longrightarrow \;
      & \tm{\cpSend{y}{z}{P}{\cpCut{x}{Q}{R}}}
        \quad \text{if} \; \notFreeIn{x}{P}
      \\
      \cpRedKappaParr
      & \tm{\cpCut{x}{\cpRecv{y}{z}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpRecv{y}{z}{\cpCut{x}{P}{R}}}
      \\
      \cpRedKappaBot
      & \tm{\cpCut{x}{\cpWait{y}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpWait{y}{\cpCut{x}{P}{R}}}
      \\
      \cpRedKappaPlus1
      & \tm{\cpCut{x}{\cpInl{y}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpInl{y}{\cpCut{x}{P}{R}}}
      \\
      \cpRedKappaPlus2
      &\tm{\cpCut{x}{\cpInr{y}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpInr{y}{\cpCut{x}{P}{R}}}
      \\
      \cpRedKappaWith
      & \tm{\cpCut{x}{\cpCase{y}{P}{Q}}{R}}
      & \Longrightarrow \;
      & \tm{\cpCase{y}{\cpCut{x}{P}{R}}{\cpCut{x}{Q}{R}}}
      \\
      \cpRedKappaTop
      & \tm{\cpCut{x}{\cpAbsurd{y}}{R}}
      & \Longrightarrow \;
      & \tm{\cpAbsurd{y}}
    \end{array}
  \]
  \begin{center}
    \begin{prooftree*}
      \AXC{$\reducesto{P}{P^\prime}$}
      \SYM{\cpRedGammaCut}
      \UIC{$\reducesto{\cpCut{x}{P}{Q}}{\cpCut{x}{P^\prime}{Q}}$}
    \end{prooftree*}
    \begin{prooftree*}
      \AXC{$\tm{P}\equiv\tm{Q}$}
      \AXC{$\reducesto{Q}{Q^\prime}$}
      \AXC{$\tm{Q^\prime}\equiv\tm{P^\prime}$}
      \SYM{\cpRedGammaEquiv}
      \TIC{$\reducesto{P}{P^\prime}$}
    \end{prooftree*}
  \end{center}
  \caption{Classical Processes, reduction relation.}
  \label{fig:cp-reduction}
\end{figure}

Reductions relate processes with their reduced forms \eg, a reduction $\reducesto{P}{Q}$ denotes that the process $\tm{P}$ can reduce to the process $\tm{Q}$ in a single step.
\begin{definition}[Reduction]\label{def:cp-reduction}
  Reductions are described by the smallest relation $\Longrightarrow$ on process terms closed under rules in \cref{fig:cp-reduction}.
  The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of 
  $\Longrightarrow$.
\end{definition}

Channels in \cp are typed using a session type system which corresponds to
classical linear logic.
A process communicating over a channel of type \ty{A \tens B} must be ready to
split into two independent processes (two processes which share no
channels), ready to communicate over a channel of type \ty{A} and \ty{B},
respectively, and process communicating over a channel of type \ty{A \parr B} is
dual to this, and therefore must be ready to start communicating concurrently
with two independent processes.
A process communicating over a channel of type \ty{A \plus B} can choose to
communicate either as \ty{A} or \ty{B}, and a process communicating over a
channel of type \ty{A \with B} must be able to communicate either as \ty{A} or
\ty{B}.
\begin{definition}[Types]\label{def:cp-types}
  \[
    \begin{array}{rclrcl}
      \ty{A}, \ty{B}, \ty{C}
           :=& \ty{A \tens B} &\text{pair of independent processes}
      &  \mid& \ty{\one}      &\text{unit for} \; {\tens}
      \\ \mid& \ty{A \parr B} &\text{pair of interdependent processes}
      &  \mid& \ty{\bot}      &\text{unit for} \; {\parr}
      \\ \mid& \ty{A \plus B} &\text{internal choice}
      &  \mid& \ty{\nil}      &\text{unit for} \; {\plus}
      \\ \mid& \ty{A \with B} &\text{external choice}
      &  \mid& \ty{\top}      &\text{unit for} \; {\with}
    \end{array}
  \]
\end{definition}
Duality plays a crucial role in both linear logic and session types.
In \cp, the two endpoints of a channel are assigned dual types.
This ensures that, for instance, whenever a process \emph{sends} across a
channel, the process on the other end of that channel is waiting to
\emph{receive}.
Each type $\ty{A}$ has a dual, written $\ty{A^\bot}$.
Duality is an involutive function.
\begin{definition}[Duality]\label{def:cp-negation}
  \[
    \begin{array}{lcllcllcllcl}
         \ty{(A \tens B)^\bot} &=& \ty{A^\bot \parr B^\bot}
      &  \ty{\one^\bot}        &=& \ty{\bot}
      &  \ty{(A \parr B)^\bot} &=& \ty{A^\bot \tens B^\bot}
      &  \ty{\bot^\bot}        &=& \ty{\one}
      \\ \ty{(A \plus B)^\bot} &=& \ty{A^\bot \with B^\bot}
      &  \ty{\nil^\bot}        &=& \ty{\top}
      &  \ty{(A \with B)^\bot} &=& \ty{A^\bot \plus B^\bot}
      &  \ty{\top^\bot}        &=& \ty{\nil}
    \end{array}
  \]
\end{definition}
Environments associate channels with types.
Names in environments must be unique, and environments $\ty{\Gamma}$ and
$\ty{\Delta}$ can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if
$\ty{\fv{\Gamma}} \cap \ty{\fv{\Delta}} = \varnothing$.
\begin{definition}[Environments]\label{def:cp-environments}
  \(
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} := \tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}
  \) 
\end{definition}
\begin{figure}
%  Structural Rules.
  \begin{center}\cpInfAx     \cpInfCut  \end{center}\vspace*{-.5\baselineskip}
%  Logical Rules.
  \begin{center}\cpInfTens   \cpInfParr \cpInfOne    \end{center}\vspace*{-.5\baselineskip}
%  \begin{center}\end{center}\vspace*{-.5\baselineskip}
  \begin{center}\cpInfBot  \cpInfPlus1  \cpInfPlus2\end{center}\vspace*{-.5\baselineskip}
  \begin{center}\cpInfWith               
%  \end{center}\vspace*{-.5\baselineskip}\begin{center}
  \cpInfNil    \cpInfTop  \end{center} %\vspace*{-.5\baselineskip}
  \caption{Classical Processes, typing.}
  \label{fig:cp}
\end{figure}
\noindent
Typing judgements associate processes with collections of typed channels.
\begin{definition}[Typing judgements]\label{def:cp}
  A typing judgement $\seq[{ P }]{\tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}}$ denotes
  that the process \tm{P} communicates along channels $\tm{x_1}\dots\tm{x_n}$
  following protocols $\ty{A_1}\dots\ty{A_n}$.
  Typing judgements are derived using rules in \cref{fig:cp}.
\end{definition}
\cp enjoys subject reduction, progress, and termination~\cite{wadler2012}.
\begin{lemma}[Preservation for $\equiv$]\label{lem:cp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\Gamma}$ iff $\seq[Q]{\Gamma}$.
\end{lemma}
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:cp-preservation}
  If $\seq[P]{\Gamma}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\Gamma}$.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}
\begin{theorem}[Progress]\label{thm:cp-progress}
  If $\seq[P]{\Gamma}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\seq[P]{\Gamma}$.
  If the last rule is \textsc{Cut}, there are four cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if either side of the cut is an axiom, we apply \cpRedAxCut1;
  \item
    if either side of the cut is itself a cut, we recursively eliminate the cut;
  \item
    if both sides are logical rules acting on the cut formula, we apply
    one of the $\beta$-rules;
  \item
    otherwise, at least one side is a logical rule acting on a formula other
    than the cut formula, in which case we apply one of the $\kappa$-rules.
  \end{enumerate*}
\end{proof}

\section{Hypersequent Classical Processes}
\label{sec:hcp}

\begin{definition}[Terms]\label{def:hcp-terms}
  \[
    \begin{array}{rllrll}
      \tm{P}, \tm{Q}, \tm{R}
           :=& \tm{\cpLink{x}{y}}         &\text{link}
      &  \mid& \tm{\piHalt}               &\text{terminated process}
      \\ \mid& \tm{\piNew{x}{P}}          &\text{name restriction, ``cut''}
      &  \mid& \tm{( \piPar{P}{Q} )}      &\text{parallel composition, ``mix''}
      \\ \mid& \tm{\piBoundSend{x}{y}{P}} &\text{``output''}
      &  \mid& \tm{\piRecv{x}{y}{P}}      &\text{``input''}
      \\ \mid& \tm{\piBoundSend{x}{}{P}}  &\text{``halt''}
      &  \mid& \tm{\cpWait{x}{}{P}}       &\text{wait}
      \\ \mid& \tm{\cpInl{x}{P}}          &\text{select left choice}
      &  \mid& \tm{\cpInr{x}{P}}          &\text{select right choice}
      \\ \mid& \tm{\cpCase{x}{P}{Q}}      &\text{offer binary choice}
      &  \mid& \tm{\cpAbsurd{x}}          &\text{offer nullary choice}
    \end{array}
  \]
\end{definition}

\begin{definition}[Structural congruence]\label{def:hcp-equiv}
  The structural congruence $\equiv$ is the congruence closure over terms which
  satisfies the following additional axioms:
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{llcllllcll}
        \cpEquivLinkComm
      & \tm{\cpLink{x}{y}}
      & \equiv
      & \tm{\cpLink{y}{x}}
      &
      &
        \hccpEquivMixHalt1
      & \tm{\piPar{P}{\piHalt}}
      & \equiv
      & \tm{P}
      &
      \\
        \hccpEquivMixComm
      & \tm{\piPar{P}{Q}}
      & \equiv
      & \tm{\piPar{Q}{P}}
      &
      &
        \hccpEquivMixAss1
      & \tm{\piPar{P}{( \piPar{Q}{R} )}}
      & \equiv
      & \tm{\piPar{( \piPar{P}{Q} )}{R}}
      &
      \\
        \hccpEquivNewComm
      & \tm{\piNew{x}{\piNew{y}{P}}}
      & \equiv
      & \tm{\piNew{y}{\piNew{x}{P}}}
      &
      &
        \hccpEquivScopeExt1
      & \tm{\piNew{x}{( \piPar{P}{Q} )}}
      & \equiv
      & \tm{\piPar{P}{\piNew{x}{Q}}}
      & \text{if }\notFreeIn{x}{P}
    \end{array}
  \]
\end{definition}

\begin{figure}[t]
  \[
    \begin{array}{llcll}
      \hccpRedAxCut1
      & \tm{\cpCut{x}{\cpLink{w}{x}}{P}}
      & \Longrightarrow
      & \tm{\cpSub{w}{x}{P}}
      \\
      \hccpRedBetaTensParr
      & \tm{\cpCut{x}{\piBoundSend{x}{y}{P}}{\piRecv{x}{y}{R}}}
      & \Longrightarrow
      & \tm{\piNew{x}{\piNew{y}{(\piPar{P}{R})}}}
      \\
      \hccpRedBetaOneBot
      & \tm{\cpCut{x}{\piBoundSend{x}{}{P}}{\piRecv{x}{}{P}}}
      & \Longrightarrow
      & \tm{\piPar{P}{Q}}
      \\
      \hccpRedBetaPlusWith1
      & \tm{\cpCut{x}{\cpInl{x}{P}}{\cpCase{x}{Q}{R}}}
      & \Longrightarrow
      & \tm{\cpCut{x}{P}{Q}}
      \\
      \hccpRedBetaPlusWith2
      & \tm{\cpCut{x}{\cpInr{x}{P}}{\cpCase{x}{Q}{R}}}
      & \Longrightarrow
      & \tm{\cpCut{x}{P}{R}}
      \\
      \hccpRedKappaNuTens
      & \tm{\piNew{x}{\piBoundSend{y}{z}{P}}}
      & \Longrightarrow
      & \tm{\piBoundSend{y}{z}{\piNew{x}{P}}}
      \\
      \hccpRedKappaNuParr
      & \tm{\piNew{x}{\piRecv{y}{z}{P}}}
      & \Longrightarrow
      & \tm{\piRecv{y}{z}{\piNew{x}{P}}}
      \\
      \hccpRedKappaNuOne
      & \tm{\piNew{x}{\piBoundSend{y}{}{P}}}
      & \Longrightarrow
      & \tm{\piBoundSend{y}{}{\piNew{x}{P}}}
      \\
      \hccpRedKappaNuBot
      & \tm{\piNew{x}{\piRecv{y}{}{P}}}
      & \Longrightarrow
      & \tm{\piRecv{y}{}{\piNew{x}{P}}}
      \\
      \hccpRedKappaNuPlus1
      & \tm{\piNew{x}{\cpInl{y}{P}}}
      & \Longrightarrow
      & \tm{\cpInl{y}{\piNew{x}{P}}}
      \\
      \hccpRedKappaNuPlus2
      & \tm{\piNew{x}{\cpInr{y}{P}}}
      & \Longrightarrow
      & \tm{\cpInr{y}{\piNew{x}{P}}}
      \\
      \hccpRedKappaNuWith
      & \tm{\piNew{x}{\cpCase{y}{P}{Q}}}
      & \Longrightarrow
      & \tm{\cpCase{y}{\piNew{x}{P}}{\piNew{x}{Q}}}
      \\
      \hccpRedKappaNuTop
      & \tm{\piNew{x}{\cpAbsurd{y}}}
      & \Longrightarrow
      & \tm{\cpAbsurd{y}}
      \\
      \hccpRedKappaMixTens
      & \tm{(\piPar{\piBoundSend{y}{z}{P}}{R})}
      & \Longrightarrow
      & \tm{\piBoundSend{y}{z}{(\piPar{P}{R})}}
      \\
      \hccpRedKappaMixParr
      & \tm{(\piPar{\piRecv{y}{z}{P}}{R})}
      & \Longrightarrow
      & \tm{\piRecv{y}{z}{(\piPar{P}{R})}}
      \\
      \hccpRedKappaMixOne
      & \tm{(\piPar{\piBoundSend{y}{}{P}}{R})}
      & \Longrightarrow
      & \tm{\piBoundSend{y}{}{(\piPar{P}{R})}}
      \\
      \hccpRedKappaMixBot
      & \tm{(\piPar{\piRecv{y}{}{P}}{R})}
      & \Longrightarrow
      & \tm{\piRecv{y}{}{(\piPar{P}{R})}}
      \\
      \hccpRedKappaMixPlus1
      & \tm{(\piPar{\cpInl{y}{P}}{R})}
      & \Longrightarrow
      & \tm{\cpInl{y}{(\piPar{P}{R})}}
      \\
      \hccpRedKappaMixPlus2
      & \tm{(\piPar{\cpInr{y}{P}}{R})}
      & \Longrightarrow
      & \tm{\cpInr{y}{(\piPar{P}{R})}}
      \\
      \hccpRedKappaMixWith
      & \tm{(\piPar{\cpCase{y}{P}{Q}}{R})}
      & \Longrightarrow
      & \tm{\cpCase{y}{(\piPar{P}{R})}{(\piPar{Q}{R})}}
      \\
      \hccpRedKappaMixTop
      & \tm{(\piPar{\cpAbsurd{y}}{R})}
      & \Longrightarrow
      & \tm{\cpAbsurd{y}}
    \end{array}
  \]
  \begin{center}
    \begin{prooftree*}
      \AXC{$\reducesto{P}{P^\prime}$}
      \SYM{\hccpRedGammaNew}
      \UIC{$\reducesto{\piNew{x}{P}}{\piNew{x}{P^\prime}}$}
    \end{prooftree*}
    \begin{prooftree*}
      \AXC{$\reducesto{P}{P^\prime}$}
      \SYM{\hccpRedGammaMix}
      \UIC{$\reducesto{\piPar{P}{Q}}{\piPar{P^\prime}{Q}}$}
    \end{prooftree*}
    \begin{prooftree*}
      \AXC{$\tm{P}\equiv\tm{Q}$}
      \AXC{$\reducesto{Q}{Q^\prime}$}
      \AXC{$\tm{Q^\prime}\equiv\tm{P^\prime}$}
      \SYM{\hccpRedGammaEquiv}
      \TIC{$\reducesto{P}{P^\prime}$}
    \end{prooftree*}
  \end{center}
  \caption{Hypersequent Classical Processes, reduction relation.}
  \label{fig:hcp-reduction}
\end{figure}

\begin{definition}[Reduction]\label{def:hcp-reduction}
 Reductions are described by the smallest relation $\Longrightarrow$ on process terms closed under rules in \cref{fig:hcp-reduction}.
%  A reduction $\reducesto{P}{Q}$ denotes that the process $\tm{P}$ can reduce
%  to the process $\tm{Q}$ in a single step.
%  The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of
%  $\Longrightarrow$.
%  Reductions can only be constructed as follows:
\end{definition}

\begin{figure}[t]
%  Structural Rules.
  \begin{center}\hccpInfAx        \hccpInfCycle\end{center}\vspace*{-.5\baselineskip}
  \begin{center}\hccpInfMix       \hccpInfHalt \end{center}\vspace*{-.5\baselineskip}
%  Logical Rules.
  \begin{center}\hccpInfBoundTens \hccpInfParr \hccpInfOne \end{center}\vspace*{-.5\baselineskip}
  \begin{center}\hccpInfBot \hccpInfPlus1 \hccpInfPlus2\end{center}\vspace*{-.5\baselineskip}
  \begin{center}\hccpInfWith \quad \hccpInfNil \quad \hccpInfTop  \end{center} %\vspace*{.5\baselineskip}
  \caption{Hypersequent Classical Processes, typing.}
  \label{fig:hcp}
\end{figure}

\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes
  that the process $\tm{P}$ consists of $n$ independent, but potentially
  interleaved processes, each of which communicates according to its own
  protocol $\Gamma_i$. Typing judgements are derived using rules in \cref{fig:hcp}---each rule has the side condition that $\tm{x} \not\in    \ty{\mathcal{G}}$.
\end{definition}

\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal G}$.
\end{lemma}
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}

\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal G}$.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}

\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\mathcal{G}}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut or a mix.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\seq[P]{\mathcal{G}}$.
  If the last rule is a cut, there are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the cut is the corresponding mix, we eliminate both as in
    \cref{thm:cp-progress};
  \item
    if the rule under the cut is an unrelated mix, we rewrite by
    \hccpEquivScopeExt{}, and recursively eliminate the cut;
  \item
    if the rule under the cut is a logical rule, we apply one of the
    $\kappa\nu$-rules.
  \end{enumerate*}
  If the last rule is mix, there are two cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the mix is a cut or another mix, we recursively eliminate it;
  \item
    if the rule under the mix is a logical rule, we apply one of the
    $\kappa{\ppar}$-rules.
  \end{enumerate*}
\end{proof}

\section{Relation between \cp and \hcp}
\label{sec:cp2hcp}

\begin{itemize}
\item Not every \hcp process is a \cp process.
\item Every \cp process is an \hcp process.
\item \hcp processes represent multiple entangled \cp processes.
\item \hcp supports the same protocols as \cp.
\end{itemize}

\begin{example}
  \(
  \seq[\cpCase{x}{(\piPar{P}{Q})}{(\piPar{P'}{Q'})}]{
    \Gamma, \Delta, \tmty{x}{A \with B}}
  \)
\end{example}

We define a translation from terms in \cp to terms in \hcp which breaks down the term constructs in \cp into their more atomic constructs in \hcp:
\begin{definition}\label{def:cp2hcp-terms}
  \[
    \begin{array}{lcllcl}
         \tm{\mtf{\cpLink{x}{y}}}
      &  := & \tm{\cpLink{x}{y}}
      &  \tm{\mtf{\cpCut{x}{P}{Q}}}
      &  := & \tm{\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}}
      \\ \tm{\mtf{\cpSend{x}{y}{P}{Q}}}
      &  := & \tm{\piBoundSend{x}{y}{(\piPar{\mtf{P}}{\mtf{Q}})}}
      &  \tm{\mtf{\cpRecv{x}{y}{P}}}
      &  := & \tm{\piRecv{x}{y}{\mtf{P}}}
      \\ \tm{\mtf{\cpHalt{x}}}
      &  := & \tm{\piBoundSend{x}{}{\piHalt}}
      &  \tm{\mtf{\cpWait{x}{P}}}
      &  := & \tm{\piRecv{x}{}{\mtf{P}}}
      \\ \tm{\mtf{\cpInl{x}{P}}}
      &  := & \tm{\cpInl{x}{\mtf{P}}}
      &  \tm{\mtf{\cpInr{x}{P}}}
      &  := & \tm{\cpInr{x}{\mtf{P}}}
      \\ \tm{\mtf{\cpCase{x}{P}{Q}}}
      &  := & \tm{\cpCase{x}{\mtf{P}}{\mtf{Q}}}
      &  \tm{\mtf{\cpAbsurd{x}}}
      &  := & \tm{\cpAbsurd{x}}
    \end{array}
  \]
\end{definition}

\begin{theorem}\label{thm:cp2hcp-typing}
  If $\seq[P]{\Gamma}$ in \cp, then $\seq[\mtf{P}]{\Gamma}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\seq[P]{\Gamma}$.
    \begin{itemize}
  \item
    Case \textsc{Cut}. We rewrite to:
    \begin{gather*}
      \begin{array}{lcl}
        \AXC{$\seq[P]{\Gamma, \tmty{x}{A}}$}
        \AXC{$\seq[Q]{\Delta, \tmty{x}{A^\bot}}$}
        \NOM{Cut}
        \BIC{$\seq[\cpCut{x}{P}{Q}]{\Gamma, \Delta}$}
        \DisplayProof
        & \Rightarrow
        & \AXC{$\seq[\mtf{P}]{\Gamma, \tmty{x}{A}}$}
          \AXC{$\seq[\mtf{Q}]{\Delta, \tmty{x}{A^\bot}}$}
          \NOM{H-Mix}
          \BIC{$\seq[\piPar{\mtf{P}}{\mtf{Q}}]{
          \Gamma, \tmty{x}{A} \hsep \Delta, \tmty{x}{A^\bot}}$}
          \NOM{H-Cut}
          \UIC{$\seq[\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}]{
          \Gamma, \Delta}$}
          \DisplayProof
      \end{array}
    \end{gather*}
  \item
    Case $(\tens)$. We rewrite as follows:
    \begin{gather*}
      \begin{array}{lcl}
        \AXC{$\seq[P]{\Gamma, \tmty{y}{A}}$}
        \AXC{$\seq[Q]{\Delta, \tmty{x}{B}}$}
        \SYM{\tens}
        \BIC{$\seq[\cpSend{x}{y}{P}{Q}]{\Gamma, \Delta, \tmty{x}{A \tens B}}$}
        \DisplayProof
        & \Rightarrow
        & \AXC{$\seq[\mtf{P}]{\Gamma, \tmty{y}{A}}$}
          \AXC{$\seq[\mtf{Q}]{\Delta, \tmty{x}{B}}$}
          \NOM{H-Mix}
          \BIC{$\seq[\piPar{\mtf{P}}{\mtf{Q}}]{
          \Gamma, \tmty{y}{A} \hsep \Delta, \tmty{x}{B}}$}
          \SYM{\tens}
          \UIC{$\seq[\piBoundSend{x}{y}{(\piPar{\mtf{P}}{\mtf{Q}})}]{
          \Gamma, \Delta, \tmty{x}{A \tens B}}$}
          \DisplayProof
      \end{array}
    \end{gather*}
  \item
    Case $(\one)$. We rewrite as follows:
    \begin{flalign*}
      \begin{array}{lcl}
        \AXC{}\SYM{\one}
        \UIC{$\seq[\cpHalt{x}]{\tmty{x}{\one}}$}
        \DisplayProof
        & \Rightarrow
        & \AXC{}
          \NOM{H-Halt}
          \UIC{$\seq[\piHalt]{\emptyhypercontext}$}
          \SYM{\one}
          \UIC{$\seq[\piBoundSend{x}{}{\piHalt}]{\tmty{x}{\one}}$}
          \DisplayProof
      \end{array}
    \end{flalign*}
  \end{itemize}
  The other cases are trivial.
\end{proof}

\begin{theorem}\label{thm:cp2hcp-equiv}
  $\tm{P}\equiv\tm{Q}$ in \cp iff $\tm{\mtf{P}}\equiv\tm{\mtf{Q}}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
% \begin{itemize}
% \item
%   Case \cpEquivCutComm.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}}
%     & \equiv & \text{by \hccpEquivMixComm}
%     \\
%     \tm{\piNew{x}{(\piPar{\mtf{Q}}{\mtf{P}})}}
%   \end{array}
%   \end{gather*}
% \item
%   Case \cpEquivCutAss1.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\mtf{P}}{\piNew{y}{(\piPar{\mtf{Q}}{\mtf{R}})}})}}
%     & \equiv & \text{%
%                by \hccpEquivNewComm, \hccpEquivMixAss1, \hccpEquivMixComm,
%                and \hccpEquivScopeExt{}}
%     \\
%     \tm{\piNew{y}{(\piPar{\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}}{\mtf{R}})}}
%   \end{array}
%   \end{gather*}
% \end{itemize}
% The other cases are trivial.
\end{proof}

\begin{theorem}\label{thm:cp2hcp-reduction}
  If $\reducesto{P}{Q}$ in \cp, then $\reducesto{\mtf{P}}{\mtf{Q}}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the the derivation of $\reducesto{P}{Q}$.
% \begin{itemize}
% \item
%   Case \cpRedBetaTensParr.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\piBoundSend{x}{y}{(\piPar{P}{Q})}}{\piRecv{x}{z}{R}})}}
%     & \Longrightarrow & \text{by \hccpRedBetaTensParr}
%     \\
%     \tm{\piNew{x}{\piNew{y}{(\piPar{(\piPar{P}{Q})}{\cpSub{y}{z}{R}})}}}
%     & \equiv & \text{%
%                by \hccpEquivNewComm, \hccpEquivMixAss1, and \hccpEquivScopeExt1}
%     \\
%     \tm{\piNew{y}{(\piPar{P}{\piNew{x}{(\piPar{Q}{\cpSub{y}{z}{R}})}})}}
%   \end{array}
%   \end{gather*}
% \item
%   Case \cpRedBetaOneBot.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\piBoundSend{x}{}{\piHalt}}{\piRecv{x}{}{\mtf{Q}}})}}
%     & \Longrightarrow & \text{by \hccpRedBetaOneBot}
%     \\
%     \tm{(\piPar{\piHalt}{\mtf{Q}})}
%     & \equiv & \text{by \hccpEquivMixHalt1}
%     \\
%     \tm{\mtf{Q}}
%   \end{array}
%   \end{gather*}
% \end{itemize}
% The other cases are trivial.
\end{proof}

\begin{theorem}\label{hcp2cp-reduction}
  If $\reducesto{\mtf{P}}{R}$ in \hcp, then there is a $\tm{Q}$ such that
  $\reducesto{P}{Q}$ in \cp and $\tm{R}\Longrightarrow^\star\tm{\mtf{Q}}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\reducesto{\mtf{P}}{R}$.
  The cases for \hccpRedAxCut1 and the $\beta$-rules are trivial.
  For the $\kappa\nu$-rules, we rewrite by the appropriate $\kappa{\ppar}$ rule,
  and vice versa.
\end{proof}

\begin{definition}\label{def:hcp-gk}
  The relation $\tm{P}\equiv_{\gamma\kappa}\tm{Q}$ is the congruence closure of
  the $\gamma$- and $\kappa$-rules for \hcp (\cref{def:hcp-reduction}).
\end{definition}

\begin{lemma}[Disentangle]\label{lem:hcp-disentangle}
  If $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ in \hcp,
  then there exist $\seq[P_1]{\Gamma_1}, \dots, \seq[P_n]{\Gamma_n}$ in \cp,
  such that
  $\tm{P}\equiv_{\gamma\kappa}\tm{(\mtf{P_1}\ppar\dots\ppar\mtf{P_n})}$.
\end{lemma}
\begin{proof}
  We repeatedly apply the $\kappa{\ppar}$-rules to move the mixes upwards.
  There are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if a mix gets stuck under a cut, it forms a \cp cut;
  \item
    if a mix gets stuck under a $(\tens)$, it forms a \cp $(\tens)$;
  \item
    otherwise, it moves all the way to the top.
  \end{enumerate*}
  Since the only way to continue after a $(\one)$ is by using a mix, all
  applications of $(\one)$ are now followed by an application of
  \textsc{H-Halt}, forming a \cp $(\one)$.
\end{proof}

We can interpret environments as types by collapsing them using a series of
pars.
\begin{definition}\label{def:bigparr}
$\ty{\bigparr(\tmty{x_1}{A_1} , \dots , \tmty{x_n}{A_n})} =  \ty{A_1 \parr \dots \parr A_n}$.
%  \begin{gather*}
%  \begin{array}{lcl}
%    \ty{\bigparr(\tmty{x_1}{A_1} , \dots , \tmty{x_n}{A_n})}
%    & = & \ty{A_1 \parr \dots \parr A_n}
%  \end{array}
%  \end{gather*}
\end{definition}

\begin{lemma}\label{lem:cp-bigparr}
  If $\seq{\Gamma}$ in \cp, then $\seq{\bigparr\Gamma}$ in \cp.
\end{lemma}
\begin{proof}
  By induction on the structure of the environment $\Gamma$.
%  \begin{itemize}
%  \item
    If $\ty{\Gamma} = \ty{A}$, then the statement is trivially true.
%  \item
    If $\ty{\Gamma} = \ty{A, B, \Gamma'}$,
    then we apply $(\parr)$ to get $\seq{A \parr B, \Gamma'}$, and apply the induction hypothesis.
%  \end{itemize}
\end{proof}

We can interpret hyper-environments as types by collapsing them using a series
of tensors. In the case of the empty hyper-environment
$\ty{\emptyhypercontext}$, we interpret this as the unit of tensor.
\begin{definition}\label{def:bigtens}
  $\ty{\bigtens(\emptyhypercontext)} = \ty{\one}$
  and 
  $\ty{\bigtens(\Gamma_1 \hsep \dots \hsep \Gamma_n)} = \ty{\bigparr\Gamma_1 \tens \dots \tens \bigparr\Gamma_n}$.
%  \begin{gather*}
%  \begin{array}{lcl}
%    \ty{\bigtens(\emptyhypercontext)}
%    & = & \ty{\one}
%    \\
%    \ty{\bigtens(\Gamma_1 \hsep \dots \hsep \Gamma_n)}
%    & = & \ty{\bigparr\Gamma_1 \tens \dots \tens \bigparr\Gamma_n}
%  \end{array}
%  \end{gather*}
\end{definition}

\begin{theorem}\label{thm:hcp2cp-bigtens}
  If $\seq{\mathcal{G}}$ in \hcp, then $\seq{\bigtens\mathcal{G}}$ in \cp.
\end{theorem}
\begin{proof}
  By case analysis on the structure of the hyper-environment $\ty{\mathcal{G}}$.
%  \begin{itemize}
%  \item
    If $\ty{\mathcal{G}} = \ty{\emptyhypercontext}$, we apply $(\one)$.
%  \item
    If $\ty{\mathcal{G}} = \ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}$,
    we apply \cref{lem:hcp-disentangle} to obtain proofs of $\seq{\Gamma_1}, \dots,
    \seq{\Gamma_n}$ in \cp, then we apply \cref{lem:cp-bigparr} to each of
    those proofs to obtain proofs of $\seq{\bigparr\Gamma_1}, \dots,
    \seq{\bigparr\Gamma_n}$, and join them using $(\tens)$ to obtain a single
    proof of $\seq{\bigtens\mathcal{G}}$ in \cp.
%  \end{itemize}
\end{proof}

\section{Related Work}
\label{sec:related-work}

Since its inception, linear logic has been described as the logic of
concurrency \cite{girard1987}.
Correspondences between the proof theory of linear logic and variants of the
\textpi-calculus emerged soon afterwards \cite{abramsky1994,bellin1994}, by
interpreting linear logic propositions as types for channels.
Later, linearity inspired also the seminal theories of linear types for the
\textpi-calculus \cite{kobayashi1999} and session types \cite{honda1998}.
Even though the two theories do not have a direct correspondence with linear
logic, the link is still strong enough that session types
can be encoded into linear types \cite{dardha2017}.

It took more than ten years for a formal correspondence between linear logic and
(a variant of) session types to emerge, with the seminal paper by \citet{caires2010}.
This inspired the development of Classical Processes by \citet{wadler2012}.

\looseness=-1
The idea of using hypersequents to capture parallelism in linear logic judgements is not
novel: \citet{carbone2018} extended the multiplicative-additive fragment
of intuitionistic linear logic with hypersequents to type global descriptions of
process communications known as choreographies.
This work is distinct from our approach in that \hcp is based on classical linear
logic and manipulates hypersequents differently: in \citet{carbone2018}, hypersequents can be formed only when sequents share
resources (\cf \textsc{H-Mix}), and resource sharing is then explicitly tracked
using an additional connection modality (which is not present in \hcp).

%\section{Conclusions}
%\label{sec:conclusions}

\clearpage
\bibliographystyle{abbrvnat}
\bibliography{main}
\end{document}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: