### 2018-10-04 doc/notes/def-cp-action.tex
merged notes and old slides

\begin{definition}[Action]\label{def:cp-action}
  A process $\tm{P}$ \emph{acts on} a channel $\tm{x}$ if it is in one of the
  following forms:
  \begin{multicols}{3}
    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
    \item $\tm{\cpLink{x}{y}}$
    \item $\tm{\cpLink{y}{x}}$
    \item $\tm{\cpSend{x}{y}{P'}{Q'}}$
    \item $\tm{\cpRecv{x}{y}{P'}}$
    \item $\tm{\cpHalt{x}}$
    \item $\tm{\cpWait{x}{P'}}$
    \item $\tm{\cpInl{x}{P'}}$
    \item $\tm{\cpInr{x}{P'}}$
    \item $\tm{\cpCase{x}{P'}{Q'}}$
    \item $\tm{\cpAbsurd{x}}$
    \end{itemize}
  \end{multicols}
  We say a process $\tm{P}$ is an \emph{action} if it acts on some channel $\tm{x}$.
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: