### 2018-10-04 doc/notes/def-cp-canonical-forms.tex

\begin{definition}[Canonical forms]\label{def:cp-canonical-forms}
  A process $\tm{P}$ is in canonical form if it is an action, or if it is of the
  form
  \[
    \tm{\cpCut{x_1}{P_1}{\dots \cpCut{x_n}{P_n}{P_{n+1}}\dots}},
  \]
  where each $\tm{P_i}$ is an action, no $\tm{P_i}$ is a link acting on a bound
  channel, and no two actions $\tm{P_i}$ and $\tm{P_j}$ act on the same bound
  channel.

  An immediate consequence of this definition is that if a process is in
  canonical form, then at least one of the actions $\tm{P_i}$ is acting on an
  free channel.
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: