### 2018-10-04 doc/notes/def-cp-envs.tex

\begin{definition}[Environments]\label{def:cp-envs}
  We define environments as follows:
  \begin{gather*}
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} ::=
    \tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}
  \end{gather*}
  The empty environment is denoted $\emptycontext$.

  Names in environments must be unique, and environments $\ty{\Gamma}$ and
  $\ty{\Delta}$ can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if
  $\text{fv}(\ty{\Gamma}) \cap \text{fv}(\ty{\Delta}) = \varnothing$.
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: