### 2018-10-04 doc/notes/def-cp-evaluation-contexts.tex

\begin{definition}[Evaluation contexts]\label{def:cp-evaluation-contexts}
  We define evaluation contexts as:
  \begin{gather*}
    \tm{E} := \tm{\hole} \mid \tm{\cpCut{x}{E}{P}} \mid \tm{\cpCut{x}{P}{E}}
  \end{gather*}
\end{definition}
\begin{definition}[Plugging]\label{def:cp-evaluation-context-plugging}
  We define plugging for evaluation contexts as:
  \begin{gather*}
    \begin{array}{ll}
      \tm{\cpPlug{\hole}{R}}            
      & := \; \tm{R}
      \\
      \tm{\cpPlug{\cpCut{x}{E}{P}}{R}}
      & := \; \tm{\cpCut{x}{\cpPlug{E}{R}}{P}}
      \\
      \tm{\cpPlug{\cpCut{x}{P}{E}}{R}}
      & := \; \tm{\cpCut{x}{P}{\cpPlug{E}{R}}}
    \end{array}
  \end{gather*}
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: