### 2018-10-04 doc/notes/def-cp-evaluation-prefixes.tex

\begin{definition}[Evaluation prefixes]\label{def:cp-evaluation-prefixes}
  We define evaluation prefixes as:
  \begin{align*}
    \tm{G}, \tm{H} := \tm{\hole} \mid \tm{\cpCut{x}{G}{H}}
  \end{align*}
  The $\tm{\hole}$ construct represents a hole.
\end{definition}
\begin{definition}[Plugging]\label{def:cp-evaluation-prefix-plugging}
  We define plugging for an evaluation prefix with $n$ holes as:
  \begin{gather*}
    \begin{array}{ll}
      \tm{\cpPlug{\hole}{R}} & := \; \tm{R} \\
      \tm{\cpPlug{\cpCut{x}{G}{H}}{R_1 \dots R_m, R_{m+1} \dots R_{n}}}
                            & := \; \tm{\cpCut{x}{\cpPlug{G}{R_1 \dots R_m}}{\cpPlug{H}{R_{m+1} \dots R_n}}}
    \end{array}
  \end{gather*}
  Note that in the second case, $\tm{G}$ is an evaluation prefix with $m$ holes,
  and $\tm{H}$ is an evaluation prefix with $(n-m)$ holes.
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: