### 2018-10-04 doc/notes/def-hccp-canonical-forms.tex

\begin{definition}[Canonical forms]\label{def:hccp-canonical-forms}
  A process $\tm{P}$ is in canonical form if it is an action, or if it is of the
  form
  \[
    \tm{\piNew{x_1\dots x_n}{(P_1\ppar\dots\ppar P_{m+n+1})}}
  \]
  where each $\tm{P_i}$ is an action, no $\tm{P_i}$ is a link acting on a bound
  channel, and no two actions $\tm{P_i}$ and $\tm{P_j}$ act on the same bound
  channel.

  An immediate consequence of this definition is that if a process is in
  canonical form, then at least $m+1$ of the actions $\tm{P_i}$ are blocked on
  free channels.
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: