### 2018-10-04 doc/notes/def-hccp-typing.tex

\begin{definition}[Typing judgements]\label{def:hccp-typing}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes
  that the process $\tm{P}$ consists of $n$ independent, but potentially
  interleaved processes, each of which communicates according to its own
  protocol $\Gamma_i$. 
  Typing judgements can be constructed using the inference rules in
  \cref{fig:hccp}.
\end{definition}
\begin{figure*}[!htb]
  Structural Rules.
  \begin{center} \hccpInfAx        \hccpInfCycle \end{center}\vspace*{1\baselineskip}
  \begin{center} \hccpInfMix       \hccpInfHalt  \end{center}\vspace*{1\baselineskip}

  Logical Rules.
  \begin{center} \hccpInfBoundTens \hccpInfParr  \end{center}\vspace*{1\baselineskip}
  \begin{center} \hccpInfOne       \hccpInfBot   \end{center}\vspace*{1\baselineskip}
  \begin{center} \hccpInfPlus1     \hccpInfPlus2 \end{center}\vspace*{1\baselineskip}
  \begin{center} \hccpInfWith                    \end{center}\vspace*{1\baselineskip}
  \begin{center} \hccpInfNil       \hccpInfTop   \end{center}\vspace*{1\baselineskip}

  \centering
  (Where each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$.)

  \caption{Hypersequent Cyclic Classical Processes with \textsc{H-Mix}}
  \label{fig:hccp}
\end{figure*}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: