### 2018-10-04 doc/notes/def-hcp-evaluation-contexts.tex

\begin{definition}[Evaluation contexts]\label{def:hcp-evaluation-contexts}
  We extend \cref{def:cp-evaluation-contexts} with the following constructs:
  \begin{align*}
    \tm{E} := \dots \mid \tm{( \piPar{E}{P} )} \mid \tm{( \piPar{P}{E} )}
  \end{align*}
\end{definition}
\begin{definition}[Plugging]\label{def:hcp-evaluation-context-plugging}
  We extend \cref{def:cp-evaluation-context-plugging} with the following cases:
  \begin{gather*}
    \begin{array}{lcl}
      \tm{\cpPlug{( \piPar{E}{P} )}{R}}
      & := & \tm{\piPar{( \cpPlug{E}{R} )}{P}}
      \\
      \tm{\cpPlug{( \piPar{P}{E} )}{R}}
      & := & \tm{\piPar{P}{( \cpPlug{E}{R} )}}
    \end{array}
  \end{gather*}
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: