### 2018-10-04 doc/notes/def-hcp-evaluation-prefixes.tex

\begin{definition}[Evaluation prefixes]\label{def:hcp-evaluation-prefixes}
  We extend \cref{def:cp-evaluation-prefixes} with the following constructs:
  \begin{align*}
    \tm{G}, \tm{H} := \dots \mid \tm{(\piPar{G}{H})}
  \end{align*}
\end{definition}
\begin{definition}[Plugging]\label{def:hcp-evaluation-prefix-plugging}
  We define plugging for an evaluation prefix with $n$ holes as:
  \begin{gather*}
    \begin{array}{l}
      \tm{\cpPlug{(\piPar{G}{H})}{R_1 \dots R_m, R_{m+1} \dots R_{n}}}
      := \; \tm{(\piPar{\cpPlug{G}{R_1 \dots R_m}}{\cpPlug{H}{R_{m+1} \dots R_n}})}
    \end{array}
  \end{gather*}
  Note that in this case, $\tm{G}$ is an evaluation prefix with $m$ holes,
  and $\tm{H}$ is an evaluation prefix with $(n-m)$ holes.
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: