### 2018-10-04 doc/notes/def-hcp-hyperenvs.tex

\begin{definition}[Hyper-environments]\label{def:hcp-hyperenvs}
  We define hyper-environments as follows:
  \begin{gather*}
    \ty{\mathcal{G}}, \ty{\mathcal{H}}
    ::= \ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}
  \end{gather*}
  The empty hyper-environment is denoted $\ty{\emptyhypercontext}$.

  Note that variable names may occur multiple times in a hyper-environment, as
  long as names within each environment are still unique.
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: