### 2018-10-04 doc/notes/def-hcp-reduction.tex

\begin{definition}[Reduction]\label{def:hcp-reduction}
  A reduction $\reducesto{P}{Q}$ denotes that the process $\tm{P}$ can reduce
  to the process $\tm{Q}$ in a single step. Reductions can only be constructed
  using the rules in \cref{def:cp-reduction} and the following addition
  congruence rule:
  \begin{prooftree}
    \AXC{$\reducesto{P}{P^\prime}$}
    \SYM{\hcpRedGammaMix}
    \UIC{$\reducesto{\piPar{P}{Q}}{\piPar{P^\prime}{Q}}$}
  \end{prooftree}
  The relations $\Longrightarrow^{?}$, $\Longrightarrow^{+}$, and
  $\Longrightarrow^\star$ are the reflexive, transitive and reflexive,
  transitive closures of $\Longrightarrow$.
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: