### 2018-10-04 doc/notes/def-safeap-equiv.tex

\begin{definition}[Structural congruence]\label{def:safeap-equiv}
  We define the structural congruence $\equiv$ as the congruence closure over
  terms which satisfies the axioms in \cref{def:ccp-equiv} and the following
  additional axioms:
  \begin{gather*}
    \begin{array}{llcll}
      \apEquivGCNew
      & \tm{\piNew{xy}{P}}
      & \equiv
      & \tm{P}
      & \text{if }\tm{x},\tm{y}\not\in\tm{P}
    \end{array}
  \end{gather*}
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: