### 2018-10-04 doc/notes/def-safeap-reduction.tex

\begin{definition}[Reduction]\label{def:safeap-reduction}
  A reduction $\reducesto{P}{Q}$ denotes that the process $\tm{P}$ can reduce
  to the process $\tm{Q}$ in a single step. Reductions can only be constructed
  using the rules in \cref{def:cccp-reduction} and the following addition
  congruence rule:
  \begin{gather*}
    \begin{array}{llcll}
      \apRedSessInit
      & \tm{\piNew{xy}{(\piPar{\piPar{\apSend{x}{z}{P}}{\apRecv{y}{w}{Q}}}{R})}}
      & \Longrightarrow
      & \tm{\piNew{xy}{\piNew{zw}{(\piPar{\piPar{P}{Q}}{R})}}}
      \\
    \end{array}
  \end{gather*}
  The relations $\Longrightarrow^{?}$, $\Longrightarrow^{+}$, and
  $\Longrightarrow^\star$ are the reflexive, the transitive, and the reflexive,
  transitive closures of $\Longrightarrow$.
\end{definition}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: