### 2018-10-04 doc/notes/def-safeap-typing.tex

\begin{definition}[Typing judgements]\label{def:ap-typing}
  A typing judgement
  $\cseq[P]{\tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}}{\mathcal{C}}$ denotes that the
  process $\tm{P}$ communicates along channels $\tm{x_1}\dots\tm{x_n}$ following
  protocols $\ty{A_1}\dots\ty{A_n}$, and that $\tm{P}$ is free from deadlocks if
  the constraint set $\cs{\mathcal{C}}$ is satisfiable.
  Typing judgements can be constructed using the inference rules in
  \cref{fig:cccp} and those in \cref{fig:safe-ap}.
\end{definition}
\begin{figure*}[!htb]
  Logical Rules.
  \begin{center} \apInfApBang  \apInfApYnot  \end{center}\vspace*{1\baselineskip}
  \begin{center} \apInfApBangC               \end{center}\vspace*{1\baselineskip}
  \begin{center} \apInfApYnotC               \end{center}\vspace*{1\baselineskip}
  \begin{center} \apInfApBangW \apInfApYnotW \end{center}

  \caption{Safe Access Points}
  \label{fig:safe-ap}
\end{figure*}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: