### 2018-10-04 doc/notes/lem-hcp-progress-beta.tex

\begin{lemma}\label{lem:hcp-progress-beta}
  If $\seq[{ \cpPlug{G}{P_1 \dots P_n} }]{ \Gamma }$, and some $\tm{P_i}$ and
  $\tm{P_j}$, on different sides of at least one cut, act on the same bound
  channel $\tm{x}$, then there exist $\tm{E}$, $\tm{E_i}$ and $\tm{E_j}$ such
  that 
  \begin{gather*}
    \tm{\cpPlug{G}{P_1 \dots P_n}} =
    \tm{\cpPlug{E}{\cpCut{x}{\cpPlug{E_i}{P_i}}{\cpPlug{E_j}{P_j}}}}.
  \end{gather*}
\end{lemma}
\begin{proof}
  As \cref{lem:cp-progress-link}.
  The two cases for
  \begin{gather*}
    \tm{\piPar{\cpPlug{G'}{P_1 \dots P_i \dots P_j \dots P_m}}{\cpPlug{G''}{P_{m+1} \dots P_n}}}\\
    \text{and}\\
    \tm{\piPar{\cpPlug{G'}{P_1 \dots P_m}}{\cpPlug{G''}{P_{m+1} \dots P_i \dots P_j \dots P_n}}}
  \end{gather*}
  are handled exactly as the cases for cuts which do not bind $\tm{x}$.
  The case where $\tm{P_i}$ and $\tm{P_j}$ are on different sides of a mix is
  excluded by the type system.
\end{proof}
%%% Local Variables:
%%% TeX-master: "main"
%%% End: