### 2019-07-13 doc/paper/eptcsdata.tex
EPTCS metadata & "revised version" subtitle

\providecommand{\firstpage}{90}
\providecommand{\publicationstatus}{T. Ehrhard, M. Fern\'andez, V. de Paiva, L. Tortora de Falco (Eds.):\\ Linearity-TLLA 2018\\
EPTCS 292, 2019, pp.\ \pageref*{FirstPage}--\pageref{LastPage},
doi:\href{http://dx.doi.org/10.4204/EPTCS.292.5}{10.4204/EPTCS.292.5}}