### 2021-09-01 doc/paper/eptcsdata.tex
erratum

\providecommand{\firstpage}{90}
%\providecommand{\publicationstatus}{T. Ehrhard, M. Fern\'andez, V. de Paiva, L. Tortora de Falco (Eds.):\\ Linearity-TLLA 2018\\ EPTCS 292, 2019, pp.\ \pageref*{FirstPage}--\pageref{LastPage}, doi:\href{http://dx.doi.org/10.4204/EPTCS.292.5}{10.4204/EPTCS.292.5}}
\providecommand{\publicationstatus}{This is a revision of a paper with the same title,\\published in EPTCS 292, 2019, pp.90--103,\\ doi:\href{http://dx.doi.org/10.4204/EPTCS.292.5}{10.4204/EPTCS.292.5}. 
This revision has been\\communicated to EPTCS on July 29, 2019.}